Fitness tracking reveals task-specific associations between memory, mental health, and physical activity

### 2022-07-16 paper/main.tex
addressing r2 comments in main text

\begin{figure}[tp]
\centering
\includegraphics[width=0.8\textwidth]{figs/corr_MH}
\caption{\textbf{Memory performance differs according to mental health
    measures.}  The reported values in the table reflect correlations
  between each behavioral measure and mental health measure.
  Only statistically reliable correlations ($p < 0.05$, corrected) are displayed. 
  We used participants' mean recall accuracy to characterize
  performance on the free recall and
  foreign language flashcards tasks, and mean precision to
  characterize performance on the naturalistic recall tasks.  We
  characterized performance on the spatial learning task using the
  (inverted and normalized)
  intercepts and slopes of linear regressions on mean estimation
  errors as a function of the number of studied shapes (also see
  Figs.~\ref{fig:immediate_behavior}, \ref{fig:delayed_behavior},
  \frDetail, \natDetail, \vocabDetail, and \spatialDetail).  Typical
  and current stress levels were measured by self report.  Mental
  health information was inferred using each participants' list of
  self-reported medications.}
\label{fig:mh_corr}
\end{figure}

We also found reliable correlations between participants' fitness and
demographic measures and their behaviors in different tasks
(for a complete report, see Fig.~\allCorr).  For example, recent
low-to-moderate-intensity (``fat burn'') 
cardiovascular activity was positively
correlated with immediate ($r = 0.38, p = 0.03$) and delayed ($r =
0.38, p = 0.029$) recall performance on the naturalistic
memory task.  Recent increases in moderate-intensity (``cardio'')
activity over the prior 7 days (relative to the preceding 30 days) was
also positively correlated with immediate naturalistic recall
performance ($r =
0.48, p = 0.003$) and immediate recall performance on the foreign language
flashcards task ($r = 0.43, p = 0.048$).  Recent high-intensity
(``peak'') activity was positively correlated with performance on the
spatial learning task ($r = 0.34, p < 0.0001$), as were recent increases
in high-intensity activity (prior 7 days versus preceding 30 days; $r
= 0.41, p = 0.01$).  Mental health indicators, such as self-reported
stress levels and medications were also associated with differences in
memory (Figs.~\ref{fig:mh_corr}, \allCorr).  For example, self-reported stress levels at the time of test
were negatively correlated with performance on the delayed memory test
for the foreign language flashcards task ($r = -0.29, p = 0.038$),
whereas participants who were medicated for anxiety and depression
tended to perform slightly (but reliably) \textit{better} on the immediate memory test for the
foreign language flashcards task ($r = 0.11, p < 0.0001$).

@@

A number of prior studies have shown that engaging in exercise can improve cognitive
and mental health~\citep{ChanEtal12b, BrisEtal02, EtniEtal06, BassSuzu17, Ragl90, MikkEtal17, TaylEtal85, DeslEtal09, Call04,
PaluSchw00, MorrEtal18, GordEtal17, MorrEtal22, HerrEtal10}.  The majority of these studies
ask participants in an ``exercise intervention'' condition (where participants engage in
a designated physical activity or training regimen) or a ``control'' condition (where participants 
do not engage in the designated activity or training) to perform cognitive tasks or undergo
mental health screening.  In other words, most primary studies treat ``physical activity'' as a binary
variable that either is or is not present for each participant.  Most prior studies also 
track or manipulate exercise over relatively short durations (typically on the order of days or weeks).  Our current work indicates
that the true relations between physical activity, cognitive performance, and mental health
may be non-monotonic and heterogeneous across activities, tasks, and mental health measures.  These
relations can also unfold over much longer timescales than have been previously identified (on the order of months; Fig.~\ref{fig:dynamics}).
However, despite the complexities of the structures of these associations, we also found that 
they were often remarkably consistent across people.  For example, as displayed in Figure~\allCorr,
many of the associations between fitness, behavioral, and mental health
measures were consistent across over 95\% of 10,000 randomly chosen subsets
of participants.

>>>>

\begin{figure}[tp]
\centering
\includegraphics[width=0.8\textwidth]{figs/corr_MH}
\caption{\textbf{Summaries of correlations between behavioral, fitness, and mental
 health measures.}  The reported values in the tables reflect correlations
  between each behavioral measure and mental health measure.
  Only statistically reliable correlations ($p < 0.05$, corrected) are displayed.
  \textbf{a. Correlations between behavioral and mental health measures.}
  We adjusted each task's behavioral measure(s) such that more positive values reflect
  better performance on the given task.  We used participants' mean recall accuracy to characterize
  performance on the free recall and
  foreign language flashcards tasks, and mean precision to
  characterize performance on the naturalistic recall tasks.  We
  characterized performance on the spatial learning task using the
  (inverted and normalized)
  intercepts and slopes of linear regressions on mean estimation
  errors as a function of the number of studied shapes (also see
  Figs.~\ref{fig:immediate_behavior}, \ref{fig:delayed_behavior},
  \frDetail, \natDetail, \vocabDetail, and \spatialDetail).  For each mental health
  measure, more positive values denote greater severity of the given measure.  Typical
  and current stress levels were measured by self report.  Mental
  health information was inferred using each participants' list of
  self-reported medications (see \textit{Methods}).  Positive correlations indicate that
  better performance on a given behavioral task is associated with more severe 
  mental health phenotypes.  \textbf{b. Correlations between fitness
  and mental health measures.}  For each fitness measure, more positive
  values denote higher observed scores (i.e., higher resting heart rate,
  more minutes of activity or time spent in each heart rate zone, or greater heart rate variability).
  The mental health measures in this panel were treated as in Panel a.
  \textbf{c. Correlations between fitness and behavioral measures.}
  Each measure reflected in this panel was treated as in Panels a and b.}
\label{fig:corr_summaries}
\end{figure}

We also found reliable correlations between participants' fitness and
demographic measures and their behaviors in different tasks
(for a complete report, see Fig.~\allCorr).  For example, recent
low-to-moderate-intensity (``fat burn'') 
cardiovascular activity was positively
correlated with immediate ($r = 0.38, p = 0.03$) and delayed ($r =
0.38, p = 0.029$) recall performance on the naturalistic
memory task.  Recent increases in moderate-intensity (``cardio'')
activity over the prior 7 days (relative to the preceding 30 days) was
also positively correlated with immediate naturalistic recall
performance ($r =
0.48, p = 0.003$) and immediate recall performance on the foreign language
flashcards task ($r = 0.43, p = 0.048$).  Recent high-intensity
(``peak'') activity was positively correlated with performance on the
spatial learning task ($r = 0.34, p < 0.0001$), as were recent increases
in high-intensity activity (prior 7 days versus preceding 30 days; $r
= 0.41, p = 0.01$).  Mental health indicators, such as self-reported
stress levels and medications were also associated with differences in
memory (Figs.~\ref{fig:corr_summaries}, \allCorr).  For example, self-reported stress levels at the time of test
were negatively correlated with performance on the delayed memory test
for the foreign language flashcards task ($r = -0.29, p = 0.038$),
whereas participants who were medicated for anxiety and depression
tended to perform slightly (but reliably) \textit{better} on the immediate memory test for the
foreign language flashcards task ($r = 0.11, p < 0.0001$).

@@

A number of prior studies have shown that engaging in exercise can improve cognitive
and mental health~\citep{ChanEtal12b, BrisEtal02, EtniEtal06, BassSuzu17, Ragl90, MikkEtal17, TaylEtal85, DeslEtal09, Call04,
PaluSchw00, MorrEtal18, GordEtal17, MorrEtal22, HerrEtal10}.  The majority of these studies
ask participants in an ``exercise intervention'' condition (where participants engage in
a designated physical activity or training regimen) or a ``control'' condition (where participants 
do not engage in the designated activity or training) to perform cognitive tasks or undergo
mental health screening.  In other words, most primary studies treat ``physical activity'' as a binary
variable that either is or is not present for each participant.  Most prior studies also 
track or manipulate exercise over relatively short durations (typically on the order of days or weeks).  Our current work indicates
that the true relations between physical activity, cognitive performance, and mental health
may be non-monotonic and heterogeneous across activities, tasks, and mental health measures.  These
relations can also unfold over much longer timescales than have been previously identified (on the order of months; Fig.~\ref{fig:dynamics}).
However, despite the complexities of the structures of these associations, we also found that 
they were often remarkably consistent across people.  For example, as displayed in Figure~\allCorr,
many of the associations between fitness, behavioral, and mental health
measures were consistent across over 97.5\% of 10,000 randomly chosen subsets
of participants.
